[]
---
output: # rmarkdown::github_document
  html_document:
  pdf_document: default
  word_document: default
title: "Assignment 3.  Working with Data Frame. Base R Style"
---

***Note*:** *This assignment practices working with Data Frame using Base R.*

***How to do it?***: 

- Open the Rmarkdown file of this assignment ([link](assignment3.Rmd)) in Rstudio. 

- Right under each question, insert  a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question. 

- `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.  

-  Publish the html file to your Githiub Page. 

***Submission***: Submit the link on Github of the assignment to Canvas under Assignment 3. 

-------

Problems
-------
 \

1. Create the following data frame

| Rank 	| Age 	| Name  	|
|:------	|:-----	|:-------	|
| 0    	| 28  	| Tom   	|
| 1    	| 34  	| Jack  	|
| 2    	| 29  	| Steve 	|
| 3    	| 42  	| Ricky 	|
```{r}
df = data.frame(Rank = c(0,1,2,3),
                Age = c(28,34,29,42),
                Name = c('Tom', 'Jack', 'Steve','Ricky'))
df
```


2. Use `read.csv` to import the Covid19 Vaccination data from WHO: [link](https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-states.csv).
```{r}
covid <-  read.csv('https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-states.csv')

```


2. Show the names of the variables in the data
```{r}
names(covid)
```


3. How many columns and rows the data have?
```{r}
str(covid)
```


4. How many missing values are there?  Show the missing values by columns.  What variable has the most number of missing values?
```{r}
sum(is.na(covid))
colSums(is.na(covid))
```


5. What is the class of the `date` column. Change the `date` columns to `date` type using the `as.Date` function. Show the new class of the `date` column. 
```{r}
class(covid$date)
covid$date = as.Date(covid$date)

```


6. Capitalize the names of all the variables
```{r}
names(covid) <- toupper(names(covid))
table(names(covid))

```


7. Find the average number of cases per day. Find the maximum cases a day. 
```{r}
mean(covid$CASES, na.rm=TRUE)
max(covid$CASES, na.rm=TRUE)

```


8. How many states are there in the data?  
```{r}
table(covid$STATE)
```


9. Create a new variable `weekdays` to store the weekday for each rows. 
```{r}
library(dplyr)
library(lubridate)
covid$weekdays <- weekdays(covid$DATE)
table(covid$weekdays)

```


10. Create the categorical variable `death2` variable taking the values as follows

- `has_death` if there is a death that day
-  `no_death` if there is no death that day
```{r}
covid$death2 <- case_when(covid$DEATHS >1 ~ 'has_death',
                       covid$DEATHS <=0 ~ 'no_death')
table(covid$death2)
```


Find the frequency and relative frequency of `no_death` and `has_death`. 

11. Find the first quartile (Q1), second quartile (Q2) and and third quartile (Q3) of the variable `death`. (Hint:  Use the `summary` function)
```{r}
summary(covid$DEATHS)
```


12. Create the categorical variable `death3` variable taking the values as follows

- `low_death` if the number of deaths smaller than the 25 percentile (Q1)

- `mid_death` if the number of deaths from Q1 to Q3

- `high_death` if the number of deaths greater than Q3
```{r}
covid$death3 <- case_when(covid$DEATHS<853.2 ~ 'low_death',
                             covid$DEATHS<4124.0 ~ 'mid_death',
                             TRUE~'high_death')
table(covid$death3)
```


13. Find the average cases in Rhode Island in 2021
```{r}
df1<-covid[(covid$STATE=='Rhode Island')&
            (covid$DATE>='2021-01-01'),]
mean(df1$CASES)
```


14. Find the median cases by weekdays in Rhode Island in 2021 
```{r}
by(df1$CASES,df1$weekdays, FUN=mean)
```


15. Compare the median cases in Rhode Island in June, July, August and September in 2021. 
```{r}
df1$months <- months(df1$DATE)
df2 <- df1[(df1$months =='June' ) |(df1$months =='July' )|(df1$months =='August' )|(df1$months =='September' ),]
by(df2$CASES,df2$months, median)
```


16.  Find your own dataset, import it and implement the following functions on the data

- head
- str
- names
- mean, min, max
- table
- is.na
- colSums
- class
- cor
- by
- ifelse
- case_when
```{r}
setwd("C:/Users/student/Downloads")
dft <-  read.csv('train.csv')
head(dft)

```
```{r}
setwd("C:/Users/student/Downloads")
dft <-  read.csv('train.csv')
str(dft)
```
```{r}
setwd("C:/Users/student/Downloads")
dft <-  read.csv('train.csv')
names(dft)
```


```{r}
setwd("C:/Users/student/Downloads")
dft <-  read.csv('train.csv')
mean(dft$Survived)
max(dft$Survived)
median(dft$Survived)
```
```{r}
setwd("C:/Users/student/Downloads")
dft <-  read.csv('train.csv')
table(dft$Survived)
```
```{r}
setwd("C:/Users/student/Downloads")
dft <-  read.csv('train.csv')
sum(is.na(dft))
```

```{r}
setwd("C:/Users/student/Downloads")
dft <-  read.csv('train.csv')
sum(is.na(dft))
colSums(is.na(dft))
```
```{r}
setwd("C:/Users/student/Downloads")
dft <-  read.csv('train.csv')
class(dft$Survived)
```
```{r}
setwd("C:/Users/student/Downloads")
dft <-  read.csv('train.csv')
cor(dft$Fare, dft$Age,  use = "complete.obs")
```

```{r}
dft1 <- dft[dft$Sex=='female',]
by(dft1$Fare, dft1$Pclass, mean)
```
```{r}
dft$Parent <- ifelse(dft$Parch!=0, 
                    'Parent','Not Parent')
table(dft$Parent)
```
```{r}
sum(is.na(dft$Age))
```

```{r}
dft$Demographic <- case_when( dft$Age<11 ~ 'Child',
                             dft$Age<20 ~'Teenager',
                             TRUE~'Adult')
table(dft$Demographic)
```


17.  In the dataset in #16, practice the follows.  You can reuse the code of 16. 
```{r}
dft$Parent <- ifelse(dft$Parch!=0, 
                    'Parent','Not Parent')
table(dft$Parent)
```
```{r}
dft$Demographic <- case_when( dft$Age<11 ~ 'Child',
                             dft$Age<20 ~'Teenager',
                             TRUE~'Adult')
table(dft$Demographic)
```


- Create a categorical variable from a continuous variable
- From one categorical variable, create a new categorical variable with fewer categories

If you do not have a data, you can use [the titanic dataset](https://www.kaggle.com/competitions/titanic/overview), which can be downloaded at [this link](../data/titanic.csv)
