
---
output: 
  html_document:
  pdf_document: default
  word_document: default
title: "Assignment 10: Predictive Modeling - Part 1"
---

***How to do it?***: 

- Open the Rmarkdown file of this assignment ([link](assignment10.Rmd)) in Rstudio. 

- Right under each **question**, insert  a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question. 

- `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.  

-  Publish the html file to your Githiub Page. 

***Submission***: Submit the link on Github of the assignment to Canvas

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```
```{r}
df2<-  read.csv('adult_census.csv')
```
```{r}
#install.packages("caret", dep="TRUE")
```

```{r}
library(gganimate)
library(ggplot2)
library(tidyverse)
library(knitr)
library(gapminder)
library(recipes)
library(rpart)
library(lubridate)
library(rlang)
```

-------

1. Use the `Adult Census Income` dataset.  We will predict the income (whether or not it is more than 50k or not) of an adult. Import the dataset.  Partition the data into 80% training and 20% testing. 
```{r}
df<-  read.csv('adult_census.csv')
df <- df %>% select(-education.num, -workclass, -fnlwgt, -occupation,-capital.gain,-capital.loss,-hours.per.week,-native.country)
df <- df %>% rename(target=income)
df <- df %>% 
  mutate(target = as.factor(target),
         race = as.factor(race),
         sex = as.factor(sex),
         education = as.factor(education),
         marital = as.factor(marital.status),
         relationship = as.factor(relationship)
         )
library(caret)
library(rpart)
set.seed(2020)
splitIndex <- createDataPartition(df$target, p = .80, 
                                  list = FALSE)
df_train <- df[ splitIndex,]
df_test <- df[-splitIndex,]
```


2. Practice Decision Tree.  Do the follows:

  - Use `rpart` package, create a decision tree with maximum depth of 3. 
```{r}
tree_model <- rpart(target ~ ., data = df_train,
                 control = rpart.control(maxdepth = 3))
```
  

  
  - Calculate the accuracy of the model on the testing data. Notice that the positive outcome here is not `1` but `>50K` or `<50K`. 
```{r}
library(gdata)
pred <- predict(tree_model, df_test, type = "class")
#Evaluate the predictions
cm <- confusionMatrix(data = pred, reference = df_test$target, positive = ">50K")
cm$overall[1]
```
  
  - Plot the tree
```{r}
library(rattle)
fancyRpartPlot(tree_model)
```
  
  
  - Plot the variable importance by the tree
```{r}
barplot(tree_model$variable.importance)
```
  
3. Create 3 more trees and compare the testing accuracy of these trees, which tree give the highest testing accuracy.
```{r}
tree_model2 <- rpart(target ~ ., data = df_train,
                 control = rpart.control(maxdepth = 4))
splitIndex2 <- createDataPartition(df$target, p = .7, 
                                  list = FALSE)
df_train2 <- df[ splitIndex2,]
df_test2 <- df[-splitIndex2,]
tree_model3 <- rpart(target ~ ., data = df_train2,
                 control = rpart.control(maxdepth = 1))
tree_model4 <- rpart(target ~ ., data = df_train2,
                 control = rpart.control(maxdepth = 2))
pred2 <- predict(tree_model2, df_test, type = "class")
cm <- confusionMatrix(data = pred2, reference = df_test$target)
cm$overall[1]
```
```{r}
pred3 <- predict(tree_model3, df_test2, type = "class")
cm <- confusionMatrix(data = pred3, reference = df_test2$target)
cm$overall[1]
```
```{r}
pred4 <- predict(tree_model4, df_test2, type = "class")
cm <- confusionMatrix(data = pred4, reference = df_test2$target)
cm$overall[1]
```

4. Practice Random Forest.  Do the follows: 

  - Use `randomForest` package, create a random forest of 1000 trees. 
```{r}
library(randomForest)
forest_model = randomForest(target ~ ., data=df_train, na.action=na.exclude,ntree = 1000)
```
  
  - Calculate the accuracy of the model on the testing data. 
```{r}
pred <- predict(forest_model, df_test, type = "class")
cm <- confusionMatrix(data = pred, reference = df_test$target)
cm$overall[1]
```
  
  - Plot the variable importance by the forest
```{r}
barplot(importance(forest_model))
```

5. Create 3 more forests and compare the testing accuracy of these forests, which forest give the highest testing accuracy.
```{r}
forest_model2 = randomForest(target ~ ., data=df_train, na.action=na.exclude,ntree = 100)
forest_model3 = randomForest(target ~ ., data=df_train,na.action=na.exclude, ntree = 500)
forest_model4 = randomForest(target ~ ., data=df_train, na.action=na.exclude,ntree = 1500)
pred2 <- predict(forest_model2, df_test, type = "class")
cm <- confusionMatrix(data = pred2, reference = df_test$target)
cm$overall[1]
```
```{r}
pred3 <- predict(forest_model3, df_test, type = "class")
cm <- confusionMatrix(data = pred3, reference = df_test$target)
cm$overall[1]
```
```{r}
pred4 <- predict(forest_model4, df_test, type = "class")
cm <- confusionMatrix(data = pred4, reference = df_test$target)
cm$overall[1]
```

6. What is the best model (in term of testing accuracy) among all models (including trees and forests) you have trained?
Forest model 2 was the best model

