
---
output: # rmarkdown::github_document
  html_document:
    theme: cerulean
    toc: true
  pdf_document: default
  word_document: default
title: "Assignment 5: Handle Missing Values"
---

***How to do it?***: 

- Open the Rmarkdown file of this assignment ([link](assignment5.Rmd)) in Rstudio. 

- Right under each **question**, insert  a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question. 

- `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.  

-  Publish the html file to your Githiub Page. 

***Submission***: Submit the link on Github of the assignment to Canvas. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```

-------

**The data:** 

This assignment works with the Adult Census Data that can be downloaded at [this link.](https://bryantstats.github.io/math421/data/adult_census_missing.csv)

-------
```{r}
library(VIM)
library(lubridate)
library(dplyr)
library(tidyverse)
```

**Questions**

1. Use `read_csv` to import the data.  Show the number of NAs for each columns.   
```{r}
setwd("C:/Users/student/Downloads")
census<-  read.csv('adult_census_missing.csv')
colSums(is.na(census))
```
```{r}

```

2. Using function `aggr` the VIM package to plot the number of NAs for each column. 
```{r}

library(VIM)
census <- na_if(census, 'Unknown')
census <- na_if(census, 'Missing')
census <- na_if(census, 'Not Available')
aggr(census) 
```


3. Find other forms of missing values. Hint: You can use the `table` function to check if there are suspicious categories (Unknown, for example) in a variable. What all other forms of missing values shown in the data. 
```{r}
table(census$workclass)
table(census$education)
table(census$martial.status)
table(census$occupation)
table(census$relationship)
table(census$native.country)
```


4. Replace all the forms of missing values found to NA
```{r}
census <- na_if(census, 'Not Known')
census <- na_if(census, '?')
```

5. Replot the number of NAs for each column. 
```{r}
table(census$workclass)
table(census$education)
table(census$martial.status)
table(census$occupation)
table(census$relationship)
table(census$native.country)
```


6. Approach 1 to handle NAs: remove all rows that have any NAs. Save the dataset after removing as a different data. The original data is unchanged (still have NAs).  How many rows left after removing?
```{r}
remove <- drop_na(census)

```


7. Approach 2 to handle NAs:  Fill all the NAs with the previous or next value. (Hint: Using `fill` function).  Save the dataset after filling as a different data. The original data is unchanged (still have NAs). 
```{r}
impute <- census %>% fill( .direction = 'updown')
```


8. Approach 3 to handle NAs: For numeric variable, replace the NAs by the median.  For categorical variables, replace the NAs to the majority. 


```{r}
median_age <- median(census$age, na.rm=TRUE)
census$age <- replace_na(census$age, median_age)
majority_class <- names(which.max(table(census$workclass)))
census$workclass <- replace_na(census$workclass, majority_class)
median_fnlwgt <- median(census$fnlwgt, na.rm=TRUE)
census$fnlwgt <- replace_na(census$fnlwgt, median_fnlwgt)
majority_edu <- names(which.max(table(census$education)))
census$education <- replace_na(census$education, majority_edu)

majority_stat <- names(which.max(table(census$marital.status)))
census$marital.status <- replace_na(census$marital.status, majority_stat)
majority_occu <- names(which.max(table(census$occupation)))
census$occupation <- replace_na(census$occupation, majority_occu)
majority_rela <- names(which.max(table(census$relationship)))
census$relationship <- replace_na(census$relationship, majority_rela)
majority_race <- names(which.max(table(census$race)))
census$race <- replace_na(census$race, majority_race)
majority_sex <- names(which.max(table(census$sex)))
census$sex <- replace_na(census$sex, majority_sex)
median_gain <- median(census$capital.gain, na.rm=TRUE)
census$capital.gain<- replace_na(census$capital.gain, median_gain)
median_loss <- median(census$capital.loss, na.rm=TRUE)
census$capital.loss<- replace_na(census$capital.loss, median_loss)
median_hrs <- median(census$hours.per.week, na.rm=TRUE)
census$hours.per.week<- replace_na(census$hours.per.week, median_hrs)
majority_country <- names(which.max(table(census$native.country)))
census$native.country <- replace_na(census$native.country, majority_country)
median_income <- median(census$income, na.rm=TRUE)
census$income<- replace_na(census$income, median_income)
```



